---
title: "LandCover example"
author: "Nathan DeMaagd"
date: "9/22/2020"
output: html_document
---





This markdown file contains all the examples from the LandCover documentation on my GitHub page. It is fully executable and will reproduce all results and figures shown in the examples.







# Initial setup


File setup

```{r file setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(42)
library(ggplot2); library(raster); library(LandCover); library(gridExtra)

```




Example data setup

```{r example data setup}


# initialize data.frame with coordinates
dat <- expand.grid(x = 1:20, y = 1:20, KEEP.OUT.ATTRS = FALSE)


# create some data: elevation, landcover, and temp/ET dependent on elevation and landcover
dat$elevation <- with(dat, 50 + 2*x + 5*y + rnorm(nrow(dat), sd = 7))
dat$landcover <- rep(c(1, 2), each = nrow(dat)/2)
dat$temp      <- with(dat, (120-0.7*(0.5*elevation + 0.3*y - 0.5*x + ifelse(landcover == 'lc1', -30, 0) + rnorm(nrow(dat)))))
dat$ET      <- with(dat, ( 0.7*(2*elevation   + 0.5*y - 1.0*x + ifelse(landcover == 'lc1', +20, 0) + rnorm(nrow(dat)))))


# create plot of variables
plot_elevation <- ggplot(data = dat) + geom_raster(aes(x = x, y=y, fill = elevation)) + scale_fill_gradientn(colors = terrain.colors(5))             + coord_equal() + labs(title = 'Elevation data')
plot_temp      <- ggplot(data = dat) + geom_raster(aes(x = x, y=y, fill = temp))      + scale_fill_gradient(low = "green", high = "darkorange")      + coord_equal() + labs(title = 'Temperature data')
plot_ET        <- ggplot(data = dat) + geom_raster(aes(x = x, y=y, fill = ET))        + scale_fill_gradient(low = "tan", high = "darkblue")          + coord_equal() + labs(title = 'ET data')
plot_landcover <- ggplot(data = dat) + geom_raster(aes(x = x, y=y, fill = as.character(landcover))) + coord_equal() + labs(title = 'Landcover data') + scale_fill_discrete(name = 'landcover')

plot_grid <- grid.arrange(plot_elevation, plot_temp, plot_ET, plot_landcover, ncol=2)

plot_grid

#ggsave(plot = plot_grid, filename = "C:/Users/nated/OneDrive - hawaii.edu/Documents/Projects/Packages/LandCover/Figures/raster_plots.png", dpi = 300, width = 6, height = 6)




# if you have rasters, you'll need to first convert to data.frames and combine

# create rasters as an example, using existing data.frames (!!!!skip this in your own data since you already have rasters!!!!)
raster_terrain   <- rasterFromXYZ(dat[c('x', 'y', 'elevation')])
raster_landcover <- rasterFromXYZ(dat[c('x', 'y', 'landcover')])
raster_temp      <- rasterFromXYZ(dat[c('x', 'y', 'temp')])
raster_ET        <- rasterFromXYZ(dat[c('x', 'y', 'ET')])

# convert rasters to data.frames
dat_terrain   <- as.data.frame(raster_terrain,   xy = TRUE)
dat_landcover <- as.data.frame(raster_landcover, xy = TRUE)
dat_temp      <- as.data.frame(raster_temp,      xy = TRUE)
dat_ET        <- as.data.frame(raster_ET,        xy = TRUE)

# merge new data.frames by coordinates
merge_func <- function(x, y)
  merge(x, y, by=c("x", "y"), all=TRUE)

dat <- Reduce(merge_func, list(dat_terrain, dat_landcover, dat_temp, dat_ET))

rm(merge_func, dat_terrain, dat_landcover, dat_temp, dat_ET)

```









# Using `gls_spatial()`

```{r using gls_spatial}


# run regression
regression_results <- gls_spatial(data = dat, landcover_varname = 'landcover', landcover_vec = c(1,2), reg_formula = ET ~ elevation + temp, error_formula = ~x+y)

# what form of error correlation was chosen for the regressions? Linear for landcover type 1 and Ratio for landcover type 2
regression_results[[1]]$modelStruct$corStruct
regression_results[[2]]$modelStruct$corStruct

# print a summary of regression from both landcover types
summary(regression_results[[1]])
summary(regression_results[[2]])


rm(list=setdiff(ls(), c('dat', 'regression_results')))

```









# Using `gls_spatial_predict()`

Before you get into changing the landcovers and predicting new values, you'll most likely want to get predicted values for your present-day dependent variable. This is because you'll want to compare the *expected* values of today to the *expected* values of the future. If you compare the *true* values of today with the *expected* values of the future, you may get some weird predicted changes that don't make ecological, biological, etc. sense. This could be because of measurement error in the given raster pixel, mis-classification, omitted variables, or any number of reasons.

```{r using gls_spatial_predict}


# get the expected values of present-day ET, using both types of landcover
dat$ET_predicted_lc1 <- gls_spatial_predict(data = dat, reg_results = regression_results, landcover = 1)
dat$ET_predicted_lc2 <- gls_spatial_predict(data = dat, reg_results = regression_results, landcover = 2)

```

